An Intrinsically Typed Solution for the List-Machine Benchmark

### 2020-06-30 papers/SBLP2020/paper.tex
Added System F work

\paragraph{Definitional intepreters}{
The use of definitional interpreters for specifying semantics dates back to Reynold's pioneer work~\cite{Reynolds72}.
Recently, the interest on such interpreters was revitalized by~\citet{Amin17} which used definitional
interpreters, implemented in Coq, to prove type soundness theorems for non-trivial typed languages like System F$_{<:}$.
Unlike our work, Amin and Rompf's formalizations does not use intrinsically-typed representations of their syntax which cluttered
their formalization with the need of dealing with ``stuck states'' in the semantics. \citet{Poulsen18} described how
definitional interpreters for imperative languages can be much concisely implemented by using intrisincally-typed syntax by
using a library for name binding using \emph{scope graphs}, which greatly simplifies the treatment of complex
binding structures. Since the list-machine benchmark was designed to address other problems than binding, our representation
using \emph{de Bruijn} indices was sufficient to implement the desired type checker and interpreter.
Other recent application of definitional interpreters was proposed by~\citet{Rouvoet20}. The main contribution
of Rouvoet et. al. was to define interpreters for linear typed languages supporting session based concurrency. In order to model
linear typing features in the interpreter, the authors have implemented monads based on a separation algebra which they
named \emph{Market}, supporting the main operations for accessing and update the store used by the intepreter.
Another work using intrisincally-typed syntax for resource control was developed by~\citet{Thiemann19}, which
implemented an interpreter for a more realistic core functional session typed calculus including recursion and
session subtyping in Agda. Thiemann modeled the semantics as an interruptible abstract machine which provides a
simple interface to a scheduler. Since our objective was formalizing the list-machine benchmark in Agda and it
does not have linearity constraints in its state manipulation, we do not need to deal with the complexities
of linear and session types as in~\cite{Rouvoet20,Thiemann19}.}

>>>>

\paragraph{Definitional intepreters}{
The use of definitional interpreters for specifying semantics dates back to Reynold's pioneer work~\cite{Reynolds72}.
Recently, the interest on such interpreters was revitalized by~\citet{Amin17} which used definitional
interpreters, implemented in Coq, to prove type soundness theorems for non-trivial typed languages like System F$_{<:}$.
Unlike our work, Amin and Rompf's formalizations does not use intrinsically-typed representations of their syntax which cluttered
their formalization with the need of dealing with ``stuck states'' in the semantics. \citet{Poulsen18} described how
definitional interpreters for imperative languages can be much concisely implemented by using intrisincally-typed syntax by
using a library for name binding using \emph{scope graphs}, which greatly simplifies the treatment of complex
binding structures. Since the list-machine benchmark was designed to address other problems than binding, our representation
using \emph{de Bruijn} indices was sufficient to implement the desired type checker and interpreter.
Other recent application of definitional interpreters was proposed by~\citet{Rouvoet20}. The main contribution
of Rouvoet et. al. was to define interpreters for linear typed languages supporting session based concurrency. In order to model
linear typing features in the interpreter, the authors have implemented monads based on a separation algebra which they
named \emph{Market}, supporting the main operations for accessing and update the store used by the intepreter.
Another work using intrisincally-typed syntax for resource control was developed by~\citet{Thiemann19}, which
implemented an interpreter for a more realistic core functional session typed calculus including recursion and
session subtyping in Agda. Thiemann modeled the semantics as an interruptible abstract machine which provides a
simple interface to a scheduler. Since our objective was formalizing the list-machine benchmark in Agda and it
does not have linearity constraints in its state manipulation, we do not need to deal with the complexities
of linear and session types as in~\cite{Rouvoet20,Thiemann19}. A formalization of System F$_{\omega\mu}$ was
the subject of Chapman et.al. work~\cite{ChapmanKNW19}, which used an intrinsically-typed representation to 
implement a normalization by evaluation for this calculus. The reason behind such formalization effort was
the verification of a core language for smart-contracts which is based on System F$_{\omega\mu}$.}
